Poincar\'e--Friedrichs constants over triangulated domains

### 2024-06-14 main.tex
manifold discussion

% Domains starshaped wrt ball

@@

A finite family of simplices $\calT$ is a \emph{simplicial complex} or \emph{triangulation}, if it satisfies the following conditions: 
(i) $\calT$ contains all the subsimplices of its members (ii) any non-empty intersection of two members of $\calT$ is a common subsimplex of each of them. 
% (i) every subsimplex of any simplex in $\calT$ is itself part of $\calT$, and (ii) the intersection between any two simplices $T, T' \in \calT$ is either empty or a shared subsimplex of both $T$ and $T'$. 
The set of $k$-dimensional simplices in $\calT$ is denoted as $\subsimplex_{{k}}(\calT)$. 
We say that a simplicial complex $\calT$ has \textit{dimension $n$} or is \textit{$n$-dimensional} if each of its simplices is a subset of an $n$-dimensional member of that complex.\footnote{We warn the reader that simplicial complexes that we call $n$-dimensional are called purely $n$-dimensional in the literature on polytopes.} 
We also write $\underlying{\calT}$ for the \textit{underlying set} of the simplicial complex $\calT$, which is defined as the union 
\[
    \underlying{\calT} := \bigcup_{ T \in \calT } T
    .
\]
We call a set \textit{triangulable} if it is the underlying set of some triangulation.

@@

We return to the theory of triangulations,
as our main objective requires some further concepts.
We are interested in simplicial complexes that triangulated manifolds. 
We put particular focus on \emph{shellable} simplicial complexes as a particularly important class of complexes. Such simplicial complexes are constructed by successively adding simplices in a particularly well-behaved manner. Patches within triangulations are examples of such shellable complexes. 
Our main references for this section are the monographs by Ziegler~\cite{ziegler2012lectures} and Kozlov~\cite{kozlov2008combinatorial}. 
\\

@@

We define an $n$-dimensional simplicial complex to be a \emph{simplicial manifold} if the underlying set $\underlying{\calT}$ is an $n$-dimensional manifold with boundary.
We recall that this means that for every $x \in \underlying{\calT}$
there exists $\epsilon > 0$ and an embedding $\phi : \Ball_\epsilon(x) \cap \underlying{\calT} \rightarrow \bbR^n$
such that $\phi(0) = 0$ and $\phi$ is an isomorphism either onto the closed unit ball $\calB = \{ x \in \bbR^n \suchthat |x| \leq 1 \}$
or onto $\{ x \in \calB \suchthat x_1 \geq 0 \}$.
Note than any simplicial complex that triangulates an $n$-dimensional manifold must be an $n$-dimensional simplicial complex.

The following special cases receive particular interest:
a \textit{simplicial $n$-ball} is any triangulation of a topological (closed) $n$-ball,
and we sometimes call this a simplicial \textit{$n$-disk}.
A \textit{simplicial $n$-sphere} is any triangulation of a topological $n$-sphere.

We gather a few helpful properties of simplicial manifolds.

\begin{lemma}\label{lemma:characterizationofmanifoldcomplexes}
    Let $\calT$ be a finite $n$-dimensional simplicial manifold. 
    \begin{itemize}
        \item Any face $F \in \Faces(\calT)$ is not contained in the boundary if and only if it is contained in exactly two $n$-simplices.
        \item Any face $F \in \Faces(\calT)$ is contained in the boundary if and only if it is contained in exactly one $n$-simplex.
        \item The simplices contained in the boundary constitute a triangulation of the boundary.
        % \item $\calT$ is non-branching, that is, every $F \in \calT$ of dimension $n-1$ is contained in at most two $n$-simplices of $\calT$. 
        \item If $S$ is an inner simplex, then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as an inner simplex
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-sphere. 
        \item If $S$ is a boundary subsimplex, then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as a boundary simplex,
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-ball.
        \item If the underlying space of $\calT$ is connected, then $\calT$ is strongly connected.
    \end{itemize}
\end{lemma}
\begin{proof}
    \begin{enumerate}
    \item 
    We will use the following fact: if $S \in \calT$ has an inner point that lies on $\partial M$, then all inner points of $S$ are on $\partial M$. 
    Since the boundary is closed, every $S \in \calT$ is either a subset of the boundary or all its inner points lie in the interior of the manifold.

@@

    If $z_F$ is an interior point of $M$, 
    then we can assume that $\mathring B_F$ is homeomorphic to an open $n$-ball and $\partial \mathring B_F$ is homeomorphic to a sphere $S$ of dimension $n-1$. 
    At the same time, $X = F \subseteq \partial \mathring B_F$ is empty if $n=1$ or homeomorphic to a sphere $S' \subseteq S$ of dimension $n-2$. 
    We see that $\partial \mathring B_F \setminus X$ has $K$ distinct connected components. 
    But by construction $\partial \mathring B_F \setminus X$ is homeomorphic to $S \setminus S'$,
    which has exactly two distinct connected components according to the Jordan-Brouwer separation theorem.
    We conclude that $F$ is contained in two $n$-simplices of $\calT$.
    If an inner face $F$ were contained in only one $n$-simplex,
    then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
    So each inner face is contained in exactly two $n$-simplices.

    If $F$ is contained in only one $n$-simplex, then it must be a boundary face. 
    If a boundary face $F$ were contained in at least two $n$-simplices,
    then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
    So each boundary face is contained in exactly one $n$-simplex.

    \item
    Clearly, the simplices of $\calT$ contained in the boundary constitute a simplicial complex. 
    If $x \in \partial M$ is a boundary simplex, then it is an inner point of some simplex $S \in \calT$.
    But then this simplex must be a boundary simplex, so the boundary simplices form a triangulation of $\partial M$.

    \item 
    % TODO structure of stars around simplices 
    Let $S$ be any simplex with midpoint $z_F$. 
    For $\delta > 0$ small enough, 
    \begin{align*}
        B_{\delta}(z_F) = \left\{ x \in M \suchthat d(x,z_F) \leq \delta \right\},
        \quad 
        S_{\delta}(z_F) = \left\{ x \in M \suchthat d(x,z_F) = \delta \right\},
    \end{align*}
    are subsets of $\patch_{\calT}(S)$.
%     For any $K \in \patch_{\calT}(S)$ that contains $S$, 
%     we can easily find a homeomorphism

%     Then $\patch_{\calT}(S)$ is star-shaped with respect to $z_F$. 
%     We let $B$
%     Any half-ray starting at $z_F$ intersects at most one vertex in $\patch_{\calT}(S)$ other than $z_F$. 
%

    % TODO stars are strongly connected 
    \item 
    We show that each vertex star is strongly connected via a short induction argument.
    Clearly, any simplicial $1$-ball and simplicial $1$-sphere is strongly connected. 
    Now, if $n \geq 1$, then any vertex star in an $(n+1)$-dimensional simplicial manifold is strongly connected if all simplicial $n$-balls and $n$-spheres are strongly connected.
    The induction argument implies that each vertex star in $\calT$ is strongly connected.

    % strongly connected
    \item 
    If the underlying space is connected, then we easily see that the underlying space of the $1$-skeleton is path-connected.
    Given $n$-simplices $S,S' \in \calT$, 
    we can thus choose a sequence of $\hat S_0=S,\hat S_1,\dots,\hat S_m=S' \in \calT$ such that $\hat S_{i} \cap \hat S_{i-1} \neq \emptyset$ for all $1 \leq i \leq m$.
    We also notice that each vertex star is strongly connected. 
    We can thus construct change the first sequence to another $S_0=S,S_1,\dots,S'=S_m \in \calT$ 
    such that $S_{i} \cap S_{i-1}$ is a face of both $S_{i}$ and $S_{i-1}$ for all $1 \leq i \leq m$.
    \end{enumerate}
\end{proof}

Suppose that $\calT$ is an $n$-dimensional simplicial manifold. 
The simplices that lie in the boundary of the underlying set are called \emph{boundary simplices}.
If $S$ lies in the boundary of the underlying set, then it lies within some boundary face, % TODO
and so we conclude that the family of boundary faces constitute a simplicial complex, which we call \textit{boundary complex}.
Any point on the boundary lies within some boundary simplex, and hence the boundary complex is a triangulation of the boundary. 
Definitions imply that the boundary complex is a simplicial manifold of dimension $n-1$.

The boundary simplices of a simplicial manifold of dimension $n$ constitute a simplicial manifold themselves, of dimension $n-1$, which triangulate the boundary of the underlying space.

% We call an $n$-dimensional simplicial complex $\calT$ \emph{non-branching} if every $F \in \calT$ of dimension $n-1$ is contained in at most two $n$-simplices of $\calT$.

@@

\begin{remark}
    Clearly, not every simplicial complex is the triangulation of some (embedded) topological manifold with or without boundary. 
    Conversely, for each dimension $n \geq 5$ there exists an $n$-dimensional compact topological manifold without a triangulation. % TODO: are these compact?
    % TODO There are computational limits to algorithmically deciding whether a simplicial complex represents a manifold. 
\end{remark}
\begin{remark}
    Whether a simplicial complex is the triangulation of a manifold is algorithmically undecidable. 
    For that reason, there is no easy combinatorial property that indicates whether a simplicial complex triangulates a manifold.
\end{remark}

% TODO: the following lemma needs only be as strong as it aids the subsequent discussion

@@

    \caption{Example of a connected $2$-dimensional simplicial manifold, which triangulates an annulus, that does neither admit a manifold-like $0$-patching nor manifold-like $1$-patching.}\label{figure:annuluscounterexample}
\end{figure}

% \color{red}
% \begin{proposition}
%     Suppose that $\calT$ is a connected $n$-dimensional simplicial complex. Then the following are equivalent:
%     \begin{itemize}
%      \item $\calT$ is manifold-like
%      \item $\calT$ has a manifold-like $0$-patching
%      % \item There exists $1 \leq k \leq n$ such that $\calT$ has a $k$-patching
%     \end{itemize}
% \end{proposition}
% \color{black}

@@

Suppose that $\calT$ is an $n$-dimensional simplicial complex and we have an enumeration of the $n$-simplices $T_{0}, T_{1}, T_{2}, \dots \in \subsimplex_{n}(\calT)$.
For any enumeration, we call 
\[
    I_m := ( \patch_{\calT}(T_{0}) \cup \patch_{\calT}(T_{1}) \cup \dots \cup \patch_{\calT}(T_{m}) ) \cap \patch_{\calT}(T_{m+1})
\]
be the $m$-th \textit{interface complex}. Note that this intersection is automatically a subcomplex of the boundary complex of the last simplex $T_{m}$.
We call the enumeration a \emph{shelling} if each interface $I_m$ is a simplicial complex of dimension $n-1$.

@@

\begin{example}
    Any simplex $T$ trivially has a shelling. The boundary complex $\partial \calT(T)$ has a shelling:
    in fact, any enumeration of the boundary faces of $T$ constitutes a shelling.
\end{example}
\begin{example}
    The standard triangulation of the $n$-dimensional cube, the Kuhn triangulation, is shellable. 
\end{example}
\begin{example}
    Not all triangulable sets admit a triangulation that is also shellable. 
    Moreover, even if a set admits a shellable triangulation, not all of its triangulations might be shellable. 
    We refer to~\cite[Example 8.9]{ziegler2012lectures} for an examples of non-shellable triangulations of cubes and tetrahedra in three dimensions. 
\end{example}

@@

\begin{lemma}
    Let $\calT$ be a $n$-dimensional simplicial manifold and $V \in \Vertices(\calT)$ be an inner vertex.
    Then $\patch_{\calT}(V)$ is shellable. 
\end{lemma}
\begin{proof}
    Without loss of generality, $V$ is the origin. 
    Any ray starting at $V$ intersects at most one vertex of $\patch_{\calT}(V)$.
    We can also assume that all boundary vertices of $\patch_{\calT}(V)$ to be unit vectors,
    which does not affect shellability. 
    The simplicial boundary complex of $\calT$ is the boundary complex of a convex polytope,
    and is thus shellable~\cite{ziegler2012lectures}. % TODO: cite exactly 
\end{proof}

\begin{lemma}
    Let $\calT$ be a $3$-dimensional simplicial manifold and $S \in \calT$.
    Then $\patch_{\calT}(S)$ is shellable. 
\end{lemma}
\begin{proof}
    The statement is clear if $S$ is an inner or boundary face of $\calT$. 
    The statement is also easily verified if $S$ is an inner or boundary edge of $\calT$.
    When $S$ is an inner vertex, then we apply the preceding lemma.

@@

\begin{lemma}
    If an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots$, then it is a simplicial manifold.
    In particular, $\patch_{\calT}(T_{0}) \cup \patch_{\calT}(T_{1}) \cup \dots \cup \patch_{\calT}(T_{m})$ is a simplicial manifold for every $m \geq 0$.
\end{lemma}
\begin{proof}    
\end{proof}

@@

>>>>

% Domains star-shaped with respect to a ball

@@

A finite family of simplices $\calT$ is a \emph{simplicial complex} or \emph{triangulation}, if it satisfies the following conditions: 
(i) $\calT$ contains all the subsimplices of its members (ii) any non-empty intersection of two members of $\calT$ is a common subsimplex of each of them. 
% (i) every subsimplex of any simplex in $\calT$ is itself part of $\calT$, and (ii) the intersection between any two simplices $T, T' \in \calT$ is either empty or a shared subsimplex of both $T$ and $T'$. 
The set of $k$-dimensional simplices in $\calT$ is denoted as $\subsimplex_{{k}}(\calT)$. 
We say that a simplicial complex $\calT$ has \textit{dimension $n$} or is \textit{$n$-dimensional} if each of its simplices is a subset of an $n$-dimensional member of that complex.\footnote{We warn the reader that simplicial complexes that we call $n$-dimensional are called purely $n$-dimensional in the literature on polytopes.} 
We also write $\underlying{\calT}$ for the \textit{underlying set} of the simplicial complex $\calT$, which is defined as the union 
\[
    \underlying{\calT} := \bigcup_{ T \in \calT } T
    .
\]
We call a set \textit{trianguable} if it is the underlying set of some triangulation.

@@

We return to the theory of triangulations,
as our main objective requires some further concepts.
We are interested in simplicial complexes that triangulated manifolds. 
We put particular focus on \emph{shellable} simplicial complexes as a particularly important class of complexes. Such simplicial complexes are constructed by successively adding simplices in a particularly well-behaved manner. Patches within triangulations are examples of such shellable complexes. 
Our main references for this section are the monographs by Kozlov~\cite{kozlov2008combinatorial} and Ziegler~\cite{ziegler2012lectures}. 
\\

@@

We define an $n$-dimensional simplicial complex to be a \emph{manifold triangulation} if the underlying set $\underlying{\calT}$ is an $n$-dimensional manifold with boundary.
We recall that this means that for every $x \in \underlying{\calT}$
there exists $\epsilon > 0$ and an embedding $\phi : \Ball_\epsilon(x) \cap \underlying{\calT} \rightarrow \bbR^n$
such that $\phi(0) = 0$ and $\phi$ is an isomorphism either onto the closed unit ball $\calB = \{ x \in \bbR^n \suchthat |x| \leq 1 \}$
or onto $\{ x \in \calB \suchthat x_1 \geq 0 \}$.
Any simplicial complex that triangulates an $n$-dimensional manifold must be an $n$-dimensional simplicial complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO: integrate into narrative 
Those simplices of the manifold triangulation that are subsets of the boundary of the underlying manifold are called \emph{boundary simplices}. 
All other simplices of the manifold triangulation are called \emph{inner simplices}. 
We will show soon that the boundary simplices of a manifold triangulation constitute a triangulation of the manifolds boundary. 
% 
%
Suppose that $\calT$ is an $n$-dimensional manifold triangulation. 
The simplices that lie in the boundary of the underlying set are called \emph{boundary simplices}.
If $S$ lies in the boundary of the underlying set, then it lies within some boundary face.
and so we conclude that the family of boundary faces constitute a simplicial complex, which we call \textit{boundary complex}.
Any point on the boundary lies within some boundary simplex, and hence the boundary complex is a triangulation of the boundary. 
Definitions imply that the boundary complex is a manifold triangulation of dimension $n-1$.
% 
%
The boundary simplices of a manifold triangulation of dimension $n$ constitute a manifold triangulation themselves, of dimension $n-1$, which triangulate the boundary of the underlying space. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The following special cases receive particular interest:
a \textit{$n$-ball triangulation} is any triangulation of a topological (closed) $n$-ball,
and we sometimes call this a \textit{$n$-disk} triangulation.
An \textit{$n$-sphere triangulation} is any triangulation of a topological $n$-sphere.

We gather a few helpful properties of manifold triangulations. 
While the reader might deem them obvious, we nevertheless include proofs.

\begin{lemma}\label{lemma:characterizationofmanifoldcomplexes}
    Let $\calT$ be a finite $n$-dimensional simplicial complex whose underlying space is a manifold $\calM$. 
    \begin{itemize}
        \item Any face $F \in \Faces(\calT)$ is not contained in the boundary if and only if it is contained in exactly two $n$-simplices.
        \item Any face $F \in \Faces(\calT)$ is contained in the boundary if and only if it is contained in exactly one $n$-simplex.
        \item The simplices contained in the boundary constitute a triangulation of the boundary.
        % \item $\calT$ is non-branching, that is, every $F \in \calT$ of dimension $n-1$ is contained in at most two $n$-simplices of $\calT$. 
        \item If $S$ is an inner simplex, then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as an inner simplex
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-sphere. 
        \item If $S$ is a boundary subsimplex, then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as a boundary simplex,
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-ball.
        \item If the underlying space of $\calT$ is connected, then $\calT$ is strongly connected.
    \end{itemize}
\end{lemma}
\begin{proof}
    \begin{enumerate}
    \item 
    We $\mathring \calM := \calM \setminus \partial \calM$ denote the interior of the manifold. 
    We will use the following fact: if $S \in \calT$ has an inner point that lies on $\partial \calM$, then all inner points of $S$ are on $\partial \calM$. 
    Since the boundary is closed, every $S \in \calT$ is either a subset of the boundary or all its inner points lie in the interior $\mathring \calM$ of the manifold.

@@

    If $z_F$ is an interior point of $\calM$, 
    then we can assume that $\mathring B_F$ is homeomorphic to an open $n$-ball and $\partial \mathring B_F$ is homeomorphic to a sphere $S$ of dimension $n-1$. 
    At the same time, $X = F \subseteq \partial \mathring B_F$ is empty if $n=1$ or homeomorphic to a sphere $S' \subseteq S$ of dimension $n-2$. 
    We see that $\partial \mathring B_F \setminus X$ has $K$ distinct connected components. 
    But by construction $\partial \mathring B_F \setminus X$ is homeomorphic to $S \setminus S'$,
    which has exactly two distinct connected components according to the Jordan-Brouwer separation theorem.
    We conclude that $F$ is contained in two $n$-simplices of $\calT$.
    If an inner face $F$ were contained in only one $n$-simplex,
    then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
    So each inner face is contained in exactly two $n$-simplices.

    If $F$ is contained in only one $n$-simplex, then it must be a face in the boundary of $\calM$. 
    If $F$ were contained in at least two $n$-simplices,
    then we could construct a homeomorphism from the $n$-ball onto itself that maps an interior point on the its boundary, which cannot be.
    So each face within $\partial \calM$ is contained in exactly one $n$-simplex.

    \item
    Clearly, the simplices of $\calT$ contained in the boundary constitute a simplicial complex. 
    If $x \in \partial \calM$ is a boundary simplex, then it is an inner point of some simplex $S \in \calT$.
    But then this simplex must be a boundary simplex, so the boundary simplices form a triangulation of $\partial \calM$.

    \item 
    Let $S$ be any simplex with vertices $v_0, v_1, \dots, v_k$, with midpoint $z_F$, and dimension $k$.
    Let $\calS := \patch_{\calT}(S)$ be its star. 
    Each $l$-dimensional simplex $T \in \calS$ that contains $S$ 
    has vertices $v_0, v_1, \dots, v_k, v_{k+1}^{S}, \dots, v_{l}^{S}$. 
    For any such simplex, we introduce a decomposition $T_{0}, \dots, T_{k}$, where each $T_{i}$ has vertices 
    \begin{align*}
        v_0, \dots, v_{i-1}, z_F, v_{i+1}, \dots, v_k, v_{k+1}^{S}, \dots, v_{l}^{S}.
    \end{align*}
    The collection $\calS'$ of these simplices and their subsimplices constitute a simplicial complex 
    that triangulates the same underlying set as $\calS$.
    Moreover, $\calS' = \patch_{\calS'}(z_F)$. 
    In particular, $z_F$ is a boundary vertex of $\calS'$ if and only if $S$ is a boundary simplex of $\calS$.

    So it remains to study the topology of vertex stars. Suppose that $S = \{z\}$ is a singleton simplex. 
    For $\delta > 0$ small enough, 
    \begin{align*}
        B_{\delta}(z) = \left\{ x \in M \suchthat d(x,z) \leq \delta \right\},
        \quad 
        S_{\delta}(z) = \left\{ x \in M \suchthat d(x,z) = \delta \right\},
    \end{align*}
    are subsets of $\patch_{\calT}(S)$.
    For each $K \in \patch_{\calT}(S)$ that contains $S$ and which has the complementary simplex $F_S$, 
    we easily find homeomorphisms $\phi_{K} : B_{\delta}(z) \cap K \rightarrow K$
    such that 
    $\phi_{K|S}$ is the identity,
    such that 
    $\phi_{K}(S_F) = S_{\delta}(z)$,
    and such that 
    for all $K, K' \in \calT$ with $K' \subseteq K$ we have $\phi_{K|K'} = \phi_{K'}$.

    If $z$ is an inner vertex of $\calS'$, 
    then the boundary simplices of $\calS'$,
    which are the boundary simplices of $\calS$, triangulate a topological sphere of dimension $n-1$.
    If $z$ is a boundary vertex of $\calS'$,
    which is the case if and only if $S$ is a boundary simplex, 
    then the underlying space of $\carapace(S)$ is homeomorphic to $S_{\delta}(z)$, 
    which is a topological ball of dimension $n-1$.

    \item 
    We show that each vertex star is strongly connected via a short induction argument.
    Clearly, any simplicial $1$-ball and simplicial $1$-sphere is strongly connected. 
    Now, if $n \geq 1$, then any vertex star in an $(n+1)$-dimensional manifold triangulation is strongly connected if all simplicial $n$-balls and $n$-spheres are strongly connected.
    The induction argument implies that each vertex star in $\calT$ is strongly connected.

    \item 
    If the underlying space is connected, then we easily see that the underlying space of the $1$-skeleton is path-connected.
    Given $n$-simplices $S,S' \in \calT$, 
    we can thus choose a sequence of $\hat S_0=S,\hat S_1,\dots,\hat S_m=S' \in \calT$ such that $\hat S_{i} \cap \hat S_{i-1} \neq \emptyset$ for all $1 \leq i \leq m$.
    We also notice that each vertex star is strongly connected. 
    We can thus construct change the first sequence to another $S_0=S,S_1,\dots,S'=S_m \in \calT$ 
    such that $S_{i} \cap S_{i-1}$ is a face of both $S_{i}$ and $S_{i-1}$ for all $1 \leq i \leq m$.
    \end{enumerate}
    All relevant results are proven. 
\end{proof}

@@

\begin{remark}
    Clearly, not every simplicial complex is the triangulation of some (embedded) topological manifold with or without boundary. 
    Whether a simplicial complex is the triangulation of a manifold is algorithmically undecidable. 
    We therefore are not in pursuit of any easy combinatorial property that indicates whether a simplicial complex (without any further specific assumptions) triangulates a manifold.

    Conversely, not all topological manifolds, even if compact, can be described as a triangulation. 
    Such manifolds appear in dimension four and higher:
    one example is the E8 manifold, which is a compact simply-connected topological $4$-manifold that is homeomorphic to the underlying set of any simplicial complex. The construction of such counterexamples typically involves an infinite procedure.
%     For each dimension $n \geq 5$ there exists an $n$-dimensional compact topological manifold without a triangulation. 
\end{remark}

@@

    \caption{Example of a connected $2$-dimensional manifold triangulation, which triangulates an annulus, that does neither admit a manifold-like $0$-patching nor manifold-like $1$-patching.}\label{figure:annuluscounterexample}
\end{figure}

@@

Suppose that $\calT$ is an $n$-dimensional simplicial complex and we have an enumeration of the $n$-simplices $T_{0}, T_{1}, T_{2}, \dots \in \subsimplex_{n}(\calT)$.
For any enumeration, we call 
\[
    I_m 
    := 
    \big( \patch_{\calT}(T_{0}) \cup \patch_{\calT}(T_{1}) \cup \dots \cup \patch_{\calT}(T_{m}) \big) 
    \cap 
    \patch_{\calT}(T_{m+1})
\]
be the $m$-th \textit{interface complex}. Note that this intersection is automatically a subcomplex of the boundary complex of the last simplex $T_{m}$.
We call the enumeration a \emph{shelling} if each interface $I_m$ is a simplicial complex of dimension $n-1$.

@@

\begin{example}
    Any simplex $T$ trivially has a shelling. The boundary complex $\partial \calT(T)$ has a shelling:
    in fact, any enumeration of the boundary faces of $T$ constitutes a shelling.
\end{example}
\begin{example}
    The standard triangulation of the $n$-dimensional cube, the Kuhn triangulation, is shellable. 
\end{example}
\begin{example}
    Not all trianguable sets admit a triangulation that is also shellable. 
    Moreover, even if a set admits a shellable triangulation, not all of its triangulations might be shellable. 
    We refer to~\cite[Example 8.9]{ziegler2012lectures} for an examples of non-shellable triangulations of cubes and tetrahedra in three dimensions. 
\end{example}

@@

\begin{lemma}
    Let $\calT$ be a $n$-dimensional manifold triangulation and $V \in \Vertices(\calT)$ be an inner vertex.
    Then $\patch_{\calT}(V)$ is shellable. 
\end{lemma}
\begin{proof}
    Without loss of generality, $V$ is the origin. 
    Any ray starting at $V$ intersects at most one vertex of $\patch_{\calT}(V)$.
    We can also assume that all boundary vertices of $\patch_{\calT}(V)$ to be unit vectors,
    which does not affect shellability. 
    The simplicial boundary complex of $\calT$ is the boundary complex of a convex polytope,
    and is thus shellable~\cite{ziegler2012lectures}. % TODO: cite exactly 
\end{proof}

\begin{lemma}
    Let $\calT$ be a $3$-dimensional manifold triangulation and $S \in \calT$.
    Then $\patch_{\calT}(S)$ is shellable. 
\end{lemma}
\begin{proof}
    The statement is clear if $S$ is an inner or boundary face of $\calT$. 
    The statement is also easily verified if $S$ is an inner or boundary edge of $\calT$.
    When $S$ is an inner vertex, then we apply the preceding lemma.

@@

\begin{lemma}
    If an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots$, then it is a manifold triangulation.
    In particular, $\patch_{\calT}(T_{0}) \cup \patch_{\calT}(T_{1}) \cup \dots \cup \patch_{\calT}(T_{m})$ is a manifold triangulation for every $m \geq 0$.
\end{lemma}
\begin{proof}    
\end{proof}

@@

\subsection{Numerical examples}

% Square 
% L-shaped 
% Slit
% Cube 
% Fechera
% Crossed bricks
